---
geometry: left=0.8in,right=0.8in,top=0.8in,bottom=0.8in
fontsize: 11pt
title: "STA 445 Assignment 1"
output: pdf_document
---

# Problem 1

Create a vector of three elements (2,4,6) and name that vector `vec_a`. Create a second vector, `vec_b`, that contains (8,10,12). Add these two vectors together and name the result `vec_c`. Print only the result `vec_c`. Do not use the print command.

```{r}
vec_a <- c(2, 4, 6)
vec_b <- c(8,10,12)
(vec_c <- vec_a + vec_b)
```


# Problem 2

Create a vector, named `vec_d`, that contains only two elements (14,20). Add this vector to `vec_a` and show the result. What is the result and what do you think R did (look up the recycling rule using Google)? What is the warning message that R gives you?

```{r message=FALSE, warning=FALSE}
vec_d <- c(14,20)
vec_d + vec_a
```
The warning given was "Warning: longer object length is not a multiple of shorter object length"
I think this is due to vec_d being a 1x2 and vec_a being a 1x3, the number of rows in d and the number of columns in a do not match.

# Problem 3

Next add 5 to the vector `vec_a` and show the result. What is the result and what did R do? Why doesn't R give you a warning message similar to what you saw in the previous problem?

```{r}
vec_a + 5
```
R added 5 to all objects in vec_a. It can do this because 5 is not a matrix, this is just regular addition.

# Problem 4

Generate the vector of integers $\left\{ 1,2,\dots5\right\}$ in two different ways. 

a.  First using the `seq()` function

```{r}
(first <- seq(1, 5, 1))
```

b. Using the `a:b` shortcut.

```{r}
(second <- 1:5)
```


# Problem 5

Generate the vector of even numbers $\left\{ 2,4,6,\dots,20\right\}$ 

a. Using the `seq()` function  

```{r}
(first_2 <- seq(2, 20, 2))
```


b. Using the a:b shortcut and some subsequent algebra. 

```{r}
(second_2 <- (1:10*2))
```


# Problem 6

Generate a vector of 21 elements that are evenly spaced between 0 and 1 using the `seq()` command and name this vector `x`. Print the vector `x`.

```{r}
x <- seq(0, 1, length.out = 21)
x
```


# Problem 7

Generate the vector $\left\{ 2,4,8,2,4,8,2,4,8\right\}$ using the `rep()` command to replicate the vector c(2,4,8). 
  
```{r}
rep_vec <- rep(c(2, 4, 8), 3)
rep_vec
```


# Problem 8
Generate the vector $\left\{ 2,2,2,2,4,4,4,4,8,8,8,8\right\}$
using the `rep()` command. You might need to check the help file for `rep()` to see all of the options that `rep()` will accept. In particular, look at the optional argument `each=`.

```{r}
vec_rep <- rep(c(2, 4, 8), each = 4)
vec_rep
```

  
# Problem 9 

In this problem, we will work with the matrix 
$$ \left[\begin{array}{ccccc}
    2 & 4 & 6 & 8 & 10\\
    12 & 14 & 16 & 18 & 20\\
    22 & 24 & 26 & 28 & 30
    \end{array}\right]$$
    
a.  Follow the prompts below to create the matrix in two ways and save the resulting matrix as `M`.

i. Create the matrix using some combination of the `seq()` and `matrix()` commands. Print `M`.

```{r}
M <- matrix(seq(2,30,2), nrow = 3, ncol = 5, byrow = TRUE)
M
```


ii. Create the same matrix by some combination of multiple `seq()` commands and either the `rbind()` or `cbind()` command. Print `M`

```{r}
row1 <- seq(2, 10, by = 2)
row2 <- seq(12, 20, by = 2)
row3 <- seq(22, 30, by = 2)

M <- rbind(row1, row2, row3)
M
```


b. Extract the second row out of `M`.

```{r}
M[2,]
```


c. Extract the element in the third row and second column of `M`.

```{r}
M[3,2]
```


# Problem 10 

The following code creates a `data.frame` and then has two different methods for removing the rows with `NA` values in the column `Grade`. Explain the difference between the two.

```{r, eval=FALSE}
df <- data.frame(name= c('Alice','Bob','Charlie','Daniel'),
                     Grade = c(6,8,NA,9))
    
df[ -which(  is.na(df$Grade) ), ]

df[  which( !is.na(df$Grade) ), ]
```
The difference is the usage of - and !, the ! refers to the opposite of what it is in use of.

# Problem 11 

Create and manipulate a list.

a.  Create a list named my.test with elements
        + x = c(4,5,6,7,8,9,10)
        + y = c(34,35,41,40,45,47,51)
        + slope = 2.82
        + p.value = 0.000131

To show you have created your list correctly, type use `str()` to show the structure of the list.

```{r}
my.test <- list(
  x      = c(4, 5, 6, 7, 8, 9, 10),
  y      = c(34, 35, 41, 40, 45, 47, 51),
  slope  = 2.82,
  p.value = 0.000131)
str(my.test)
```

b.  Extract the second element in the list.

```{r}
my.test[2]
```


c.  Extract the element named `p.value` from the list.

```{r}
my.test["p.value"]
```


